---
title: "Data Importing - Week 5"
author: "Samed Jain"
date: "Nov 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## 11.2.2 EXERCISES

#### 1. What function would you use to read a file where fields were separated with "|"?

    read_delim(file = ..., delim = "|")
    
#### 2. Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?

    col_names, col_types, locale, na, quoted_na, trim_ws, n_max, guess_max, progress
    
#### 3. What are the most important arguments to read_fwf()?
    
    most important argument would be col_positions. it tells us where data ends and begins.
    
#### 4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you'll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?

```{r}
x <- "x,y\n1,'a,b'"
read_delim(x, delim = ",",quote = "'")
```
OR

```{r}
x <- "x,y\n1,'a,b'"
read_csv(x, quote = "'")
```

#### 5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6") # Only 2 columns specified in header, but 3 columns - so the last column gets dropped
read_csv("a,b,c\n1,2\n1,2,3,4") # number of rows and columns don't match
read_csv("a,b\n\"1", col_names = FALSE) # The opening quote `"1` is dropped because it is not closed, and `a` is treated as an integer.
read_csv("a,b\n1,2\na,b") # col_names need to be set to FALSE
read_csv("a;b\n1;3") # Change delim to ','
```

## 11.3.5 Exercises

#### 1. What are the most important arguments to locale()?

   Most important arguments are based on: date and time formats (date_names), time zone (tz), numbers (decimal_mark), encoding (encoding)
    
    
#### 2. What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default value of grouping_mark when you set decimal_mark to ","? What happens to the default value of decimal_mark when you set the grouping_mark to "."?

The error because `decimal_mark` and `grouping_mark` are not different - error will be displayed
The default value becomes "." as the "'" is replaced with "." The default value becomes "," as "." is replaced with "'"
  
#### 3. I didn't discuss the date_format and time_format options to locale(). What do they do? Construct an example that shows when they might be useful.

Used for default date and time format, more practical to use if date and time format remain constant.

```{r}
parse_date("15 janvier 2015", "%d %B %Y", locale = locale("fr"))
parse_date("20 oct. 1990", "%d %b %Y", locale = locale("fr"))
```

#### 4. If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.

```{r}
locale(date_format = "%d-%m-%Y", time_format = "%H:%M:%S", decimal_mark = ".", grouping_mark = ",")
```
#### 5. What's the difference between read_csv() and read_csv2()?

  Primary difference is that read_csv() will use a comma or period fo seperation, while the other function uses semi colon.	

#### 6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.
European 
Western European Latin script languages: ISO-8859-1, Windows-1250 (also CP-1250 for code-point)
Eastern European Latin script languages: ISO-8859-2, Windows-1252
Greek: ISO-8859-7
Turkish: ISO-8859-9, Windows-1254
Hebrew: ISO-8859-8, IBM424, Windows 1255
Russian: Windows 1251
Japanese: Shift JIS, ISO-2022-JP, EUC-JP
Korean: ISO-2022-KR, EUC-KR
Chinese: GB18030, ISO-2022-CN (Simplified), Big5 (Traditional)
Arabic: ISO-8859-6, IBM420, Windows 1256


#### 7. Generate the correct format string to parse each of the following dates and times:

```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
parse_date(d1, "%B %d, %Y")
parse_date(d2, "%Y-%b-%d")
parse_date(d3, "%d-%b-%Y")
parse_date(d4, "%B %d (%Y)")
parse_date(d5, "%m/%d/%y")
parse_time(t1, "%H%M")
```