---
title: "Data Frame and Matrices - Week 6"
author: "Samed Jain"
date: "Dec 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Chapter 8: Matrices and Dataframes from YaRrr! The Pirate’s Guide to R	

The following table shows the results of a survey of 10 pirates. In addition to some basic demographic information, the survey asked each pirate "What is your favorite superhero?"" and "How many tattoos do you have?""

```{r echo = FALSE}
superhero <- data.frame(
  Name = c("Astrid", "Lea", "Sarina", "Remon", "Letizia", "Babice", "Jonas", "Wendy", "Niveditha", "Gioia"),
  Sex = c("F", "F", "F", "M", "F", "F", "M", "F", "F", "F"),
  Age = c(30, 25, 25, 29, 22, 22, 35, 19, 32, 21),
  Superhero = c("Batman", "Superman", "Batman", "Spiderman", "Batman",
               "Antman", "Batman", "Superman", "Maggott", "Superman"),
  Tattoos = c(11, 15, 12, 5, 65, 3, 9, 13, 900, 0)
)
knitr::kable(superhero)
## Creating the  dataframe
```




1. Combine the data into a single dataframe. Complete all the following exercises from the dataframe!
```{r}
prt <- data.frame(
  name = c("Astrid", "Lea", "Sarina", "Remon", "Letizia", "Babice", "Jonas", "Wendy", "Niveditha", "Gioia"),
  sex = c("F", "F", "F", "M", "F", "F", "M", "F", "F", "F"),
  age = c(30, 25, 25, 29, 22, 22, 35, 19, 32, 21),
  superhero = c("Batman", "Superman", "Batman", "Spiderman", "Batman",
               "Antman", "Batman", "Superman", "Maggott", "Superman"),
  tattoos = c(11, 15, 12, 5, 65, 3, 9, 13, 900, 0),
  stringsAsFactors = FALSE
) 
``````

2. What is the median age of the 10 pirates?
  ##median function is used to calculate the median
```{r}
median(prt $age)
```

  
3. What was the mean age of female and male pirates separately?
## There are various ways we can do this one with mean function, with function, subset
```{r}
mean(prt$age[prt$sex == "F"])
mean(prt $age[prt$sex == "M"])
```

  
4. What was the most number of tattoos owned by a male pirate?
## We can do it by the following two ways 
```{r}
     max(tattoos[sex == "M"]))
```


5. What percent of pirates under the age of 25 were female?

```{r}
mean(sex[age < 25] == "F"))
```

6. What percent of female pirates are under the age of 25?
```{r}
mean(age[sex == "F"] < 25))
```

7. Add a new column to the dataframe called `tattoos.per.year` which shows how many tattoos each pirate has for each year in their life.

```{r}
prt$tattoos.per.year <- with(prt, tattoos / age)
```

8. Which pirate had the most number of tattoos per year?

```{r}
## Finding the max by max function
prt$name[prt$tattoos.per.year == max(prt$tattoos.per.year)]
```

9. What are the names of the female pirates whose favorite piratesurvey is Superman?
  
```{r}
prt$name[with(prt, sex == "F" & superhero == "Superman")]
```
  
10. What was the median number of tattoos of pirates over the age of 20 whose favorite piratesurvey is Spiderman?

```{r}
with(prt, (tattoos[age > 20 & superhero == "Spiderman"]))
```